---
title: "bacteria_processing_1"
author: "Ilya"
date: "1/28/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####install and load required packages
```{r packages, echo=FALSE}
source("packages_bacteria1.R")
```


###1. Get bacteria-caused dx in GIDEON   
###2. Match dx to pathogen spp. names
###3. Match spp. names (from GIDEON zdx and GMPD) to traits (in GMPD & other)
###4. Compile master list of bacteria spp & traits
###5. Feature construction with bacterial traits
###6. Data visualization: summary “state of knowledge” on bacteria causing disease in mammals or humans
###7. Use traits to predict transmissibility and human disease outcomes

###2. Match dx to pathogen spp. names
####Include non-zoonotic pathogenic bacteria from GMPD

####get zoonotic and non-zoonotic pathogens from GMPD and assign taxonomic ID only for those that are species-level
select only bacteria using info in G_taxonomy
```{r R_name2taxid_GMPD}
source("R_name2taxid_GMPD.R")
# source(paste0(wd,"/R_name2taxid_GMPD.R"))
# ../CODE/DATA_PROCESSING/
```

####Remove records that are NA for tax_id. classify pathogens in GMPD to order, family, genus, species (official species name may be different from entered name). Make G_classified.Rdata and save 
```{r R_classify_bacteria_observed_GMPD}
source("R_classify_bacteria_observed_GMPD.R")
```


####label pathogens as zoonotic, human only, or wild only. output: bacteria_pathogenic_mammals. Input: G_classified, df_all.Rdata
```{r R_human_zoonotic_non_zoonotic}
source("R_human_zoonotic_non_zoonotic1.R")
```

####check that all pathogenic bacteria are present in list of all bacteria. Output:  bacteria_species_out2.Rdata (put this on dropbox)
```{r}
source("R_graph_bacteria_order_pathogen_status2.R")
```



