---
title: "bacteria_graph_1"
author: "Ilya"
date: "1/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####install and load required packages
```{r packages, echo=FALSE}
source("packages_bacteria2.R")
```

###1. Get bacteria-caused dx in GIDEON   
###2. Match dx to pathogen spp. names
###3. Match spp. names (from GIDEON zdx and GMPD) to traits (in GMPD & other)
###4. Compile master list of bacteria spp & traits
###5. Feature construction with bacterial traits
###6. Data visualization: summary “state of knowledge” on bacteria causing disease in mammals or humans
###7. Use traits to predict transmissibility and human disease outcomes

##6. Data visualization: summary “state of knowledge” on bacteria causing disease in mammals or humans

###Processing data
####load bacteria_pathogenic_mammals.Rdata and make graph of bacteria order vs. host order
```{r}
source("R_graph_pathogen_order_host_order1.R")
plot
```

####make graph of histogram of number of hosts per bacteria, across zoonotic categories (human only, zoonotic, wild only)
```{r}
source("R_bacteria_host_species_hist_graph1.R")
plot
```




